---
title: 'BIOS 6301: Assignment 3'
author: "Charles Rhea"
output:
  pdf_document: default
  html_document: default
---

*Due Tuesday, 26 September, 1:00 PM*

50 points total.

Add your name as `author` to the file's metadata section.

Submit a single knitr file (named `homework3.rmd`) by email to marisa.h.blackman@vanderbilt.edu.
Place your R code in between the appropriate chunks for each question.
Check your output by using the `Knit HTML` button in RStudio.

$5^{n=day}$ points taken off for each day late.

### Question 1 ###

**15 points**

Write a simulation to calculate the power for the following study
design.  The study has two variables, treatment group and outcome.
There are two treatment groups (0, 1) and they should be assigned
randomly with equal probability.  The outcome should be a random normal
variable with a mean of 60 and standard deviation of 20.  If a patient
is in the treatment group, add 5 to the outcome.  5 is the true
treatment effect.  Create a linear model for the outcome by the
treatment group, and extract the p-value (hint: see assignment1).
Test if the p-value is less than or equal to the alpha level, which
should be set to 0.05.

*The p-value is greater than 0.05*

```{r}
n <- 25
treatment <- rbinom(n, size = 1, prob = 0.5)
outcome <- rnorm(n, 60, 20)
outcome_2 <- ifelse(treatment == 1, outcome+5, outcome)

model <- lm(outcome_2 ~ treatment)
summary(model)
```

```{r}
model.c <- coef(summary(model))
p <- (pt(model.c[2,3],8,lower.tail = FALSE))*2
p
```

```{r}
p < 0.05
```

Repeat this procedure 1000 times. The power is calculated by finding
the percentage of times the p-value is less than or equal to the alpha
level.  Use the `set.seed` command so that the professor can reproduce
your results.

1. Find the power when the sample size is 100 patients. (10 points)
```{r}
set.seed(1)

n <- 100
mean(replicate(1000, {
  treatment <- rbinom(n, size = 1, prob = 0.5)
  outcome <- rnorm(n, 60, 20)
  outcome_2 <- ifelse(treatment == 1, outcome+5, outcome)
  
  model <- lm(outcome_2 ~ treatment)
  model.c <- coef(summary(model))
  p <- (pt(model.c[2,3],8,lower.tail = FALSE))*2
}) < 0.05 )
```

1. Find the power when the sample size is 1000 patients. (5 points)
```{r}
set.seed(2)

n <- 1000
mean(replicate(1000, {
  treatment <- rbinom(n, size = 1, prob = 0.5)
  outcome <- rnorm(n, 60, 20)
  outcome_2 <- ifelse(treatment == 1, outcome+5, outcome)
  
  model <- lm(outcome_2 ~ treatment)
  model.c <- coef(summary(model))
  p <- (pt(model.c[2,3],8,lower.tail = FALSE))*2
}) < 0.05 )
```

### Question 2 ###

**14 points**

Obtain a copy of the [football-values lecture](https://github.com/couthcommander/football-values).
Save the `2023/proj_wr23.csv` file in your working directory.  Read
in the data set and remove the first two columns.
```{r}
library(readr)
df <- read_csv("~/Desktop/BIOS 6301 - Introduction to Statistical Computing/datasets//proj_wr23.csv")
head(df)
```

```{r}
wr23.df <- df[-c(1,2)]
head(wr23.df)
```

1. Show the correlation matrix of this data set. (4 points)
```{r}
wr23.corr <- cor(wr23.df)
wr23.corr
```

1. Generate a data set with 30 rows that has a similar correlation
structure.  Repeat the procedure 1,000 times and return the mean
correlation matrix. (10 points)
```{r}
library(MASS)

var.wr23.df <- var(wr23.df) 
mean.wr23.df <- colMeans(wr23.df)

sim <- mvrnorm(30, mu = mean.wr23.df, Sigma = var.wr23.df)
sim.df <- as.data.frame(sim)

sim.matrix <- cor(sim.df)
sim.matrix
```

```{r}
keep.1=0
loops=1000

for (i in 1:loops){
      sim2 <- mvrnorm(30, mu = mean.wr23.df, Sigma = var.wr23.df)
      sim2.matrix <- keep.1+cor(sim2)/loops
}

sim2.matrix
```

### Question 3 ###

**21 points**

Here's some code:

```{r}
nDist <- function(n = 100) {
    df <- 10
    prob <- 1/3
    shape <- 1
    size <- 16
    list(
        beta = rbeta(n, shape1 = 5, shape2 = 45),
        binomial = rbinom(n, size, prob),
        chisquared = rchisq(n, df),
        exponential = rexp(n),
        f = rf(n, df1 = 11, df2 = 17),
        gamma = rgamma(n, shape),
        geometric = rgeom(n, prob),
        hypergeometric = rhyper(n, m = 50, n = 100, k = 8),
        lognormal = rlnorm(n),
        negbinomial = rnbinom(n, size, prob),
        normal = rnorm(n),
        poisson = rpois(n, lambda = 25),
        t = rt(n, df),
        uniform = runif(n),
        weibull = rweibull(n, shape)
    )
}
```

1. What does this do? (3 points)

    ```{r}
    round(sapply(nDist(500), mean), 2)
    ```
*Here, 'sapply' is running the nDist function (defined above) on the value of 500. Further, we have specified that 'sapply' return the mean values of each output (e.g., beta, binomial, chisquared) and round the numerical value to 2 decimal places.*

1. What about this? (3 points)

    ```{r}
    sort(apply(replicate(20, round(sapply(nDist(10000), mean), 2)), 1, sd))
    ```
*We are using 'sapply' to run the nDist function on the value of 10000 and specifying it return the mean values of each output rounded to 2 decimal places. Additionally, we are asking this command to be replicated 20 times and to sort the outputted means in a matrix. Finally, we are asking to find the standard deviation of each mean, then sort the results from lowest to highest.*

    In the output above, a small value would indicate that `N=10,000` would provide a sufficent sample size as to estimate the mean of the distribution. Let's say that a value *less than 0.02* is "close enough".

1. For each distribution, estimate the sample size required to simulate the distribution's mean. (15 points)
```{r}
#sizeDist = c()
#n=100000
# 
#   while(length(sizeDist) != 15){
#    sdDist = rowMeans(replicate(100, {
#                                  apply(replicate(20, round(sapply(nDist(n), mean), 2)), 1, sd)
#                                      }))
#    x = sdDist[which(sdDist < 0.02)]
#    sizeDist[setdiff(names(x), names(sizeDist))]= n
#    n=n+1000
#  }

#sizeDist
```

Don't worry about being exact. It should already be clear that N < 10,000 for many of the distributions. You don't have to show your work. Put your answer to the right of the vertical bars (`|`) below.

distribution|N
---|---
beta| 100
binomial| 8,000
chisquared| 50,000
exponential| 2,500
f| 1,200
gamma| 2,500
geometric| 15,000
hypergeometric| 4,300
lognormal| 12,000
negbinomial| 200,000
normal| 2,500
poisson| 70,000
t| 3,000
uniform| 300
weibull| 2,600
