---
author: 'Charlie Rhea'
title: 'BIOS 6301: Assignment 8'
output: pdf_document
---

*Due Tuesday, 14 November, 1:00 PM*

$5^{n=day}$ points taken off for each day late.

30 points total.

Submit a single knitr file (named `homework8.rmd`), along with a valid PDF output file. Inside the file, clearly indicate which parts of your responses go with which problems (you may use the original homework document as a template). Add your name as `author` to the file's metadata section. Raw R code/output or word processor files are not acceptable.

Failure to name file `homework8.rmd` or include author name may result in 5 points taken off.

### Question 1 ###

**15 points**

Install the `readxl` package and run the following

```{r}
library(readxl)
fn <- 'icd10.xlsx'
if(file.access(fn, mode = 4) == -1) {
    url <- "https://www.cdc.gov/nhsn/xls/icd10-pcs-pcm-nhsn-opc.xlsx"
    download.file(url, destfile = fn, mode = 'wb')
}
dat <- readxl::read_excel(fn, sheet = 2)
```

1. Show the class of `dat`. (1 point)

*Object 'dat' has 3 classes: 'tbl_df', 'tbl', and 'data.frame'.*
```{r}
class(dat)
```

2. Show the methods available for objects of the given class (if there are multiple classes, show methods for all classes). (3 points)

```{r}
methods(class="tbl_df")
```

```{r}
methods(class="tbl")
```

```{r}
methods(class="data.frame")
```

3. If you call `print(dat)`, what print method is being dispatched? (1 point)

*There is a 'print.tbl_df' method within print. Since we know 'tbl_df' is the first class type for 'dat', then 'print.tbl_df' is the method that is being dispatched when we call print(dat).*
```{r}
#methods(print)
print(head(dat))
```

4. Set the class of `dat` to be a data.frame. (1 point)
```{r}
dat <- as.data.frame(dat)
class(dat)
```

5. If you call `print(dat)` again, what print method is being dispatched? (1 point)

*Since 'dat' is now a data.frame, when we call the print commend, the print.data.frame method is being dispatched*
```{r}
print(head(dat))
```

Define a new generic function `nUnique` with the code below.
```{r}
nUnique <- function(x) {
    UseMethod('nUnique')
}
```

6. Write a default method for `nUnique` to count the number of unique values in an element. (2 points)

*Using the nUnique default method, I was able to determine the number of unique values for two variables within the 'dat' dataset. For 'Procedure Code Category' I found 39 unique values, and for 'ICD-10-PCS Codes' I found 8993 unique values.*
```{r}
nUnique.default <- function(x) {
    length(unique(x))
}

v1 <- dat$`Procedure Code Category`
nUnique(v1)
```

```{r}
v2 <- dat$`ICD-10-PCS Codes`
nUnique(v2)
```

7. Check your function (2 points)
```{r, eval = FALSE}
nUnique(letters) # should return 26
```

```{r}
nUnique(sample(10, 100, replace = TRUE)) # should return 10 (probably)
```

8. Write a data.frame method for `nUnique` to operate on data.frame objects.
This version should return counts for each column in a data.frame. (2 points)
```{r}
nUnique.data.frame <- function(x) {
  for(i in colnames(x)){
    cat("Unique values in", i, ":", length(unique(x[,i])), "\n")
}}
```

9. Check your function (2 points)
```{r, eval = FALSE}
nUnique(dat)
```

### Question 2 ###

**15 points**

Programming with classes. The following function will generate random patient information.
```{r}
makePatient <- function() {
  vowel <- grep("[aeiou]", letters)
  cons <- grep("[^aeiou]", letters)
  name <- paste(sample(LETTERS[cons], 1), sample(letters[vowel], 1), sample(letters[cons], 1), sep='')
  gender <- factor(sample(0:1, 1), levels=0:1, labels=c('female','male'))
  dob <- as.Date(sample(7500, 1), origin="1970-01-01")
  n <- sample(6, 1)
  doa <- as.Date(sample(1500, n), origin="2010-01-01")
  pulse <- round(rnorm(n, 80, 10))
  temp <- round(rnorm(n, 98.4, 0.3), 2)
  fluid <- round(runif(n), 2)
  list(name, gender, dob, doa, pulse, temp, fluid)
}
```

1. Create an S3 class `medicalRecord` for objects that are a list with the named elements `name`, `gender`, `date_of_birth`, `date_of_admission`, `pulse`, `temperature`, `fluid_intake`. Note that an individual patient may have multiple measurements for some measurements.  Set the RNG seed to `8` and create a medical record by taking the output of `makePatient`.  Print the medical record, and print the class of the medical record. (5 points)
```{r}
set.seed(8)
p <- makePatient()
names(p) <- c("name", "gender", "date_of_birth", "date_of_admission", "pulse", "temperature", "fluid_intake")
class(p) <- 'medicalRecord'
print(p)
```

2. Write a `medicalRecord` method for the generic function `mean`, which returns averages for pulse, temperature and fluids. Also write a `medicalRecord` method for `print`, which employs some nice formatting, perhaps arranging measurements by date, and `plot`, that generates a composite plot of measurements over time.  Call each function for the medical record created in part 1. (5 points)
points)
```{r}
mean.medicalRecord <- function(x) {
  list(mean(p[[5]]),mean(p[[6]]), mean(p[[7]]))
}

p2 <- mean(p)
names(p2) <- c("Avg. Pulse", "Avg. Temperature", " Avg. Fluid Intake")
print(p2)
```

```{r}
print.medicalRecord<- function(med) {
  cat(sprintf("Date of Admission: %s\nName: %s\nGender: %s\nDate of Birth: %s\nPulse: %s\nTemperature: %s\nFluid Intake %s\n", 
med$date_of_admission, med$name, med$gender, med$date_of_birth, med$pulse, med$temperature, med$fluid_intake), "\n")
}

print(p)
```

```{r}
plot.medicalRecord <- function(x){
plot(x[[5]],type='b',col='red',main="Change in Pulse by Admission",xlab='Admission #', ylab='Pulse Value')
plot(x[[6]],type='b',col='blue',main="Change in Temperature by Admission",xlab='Admission #', ylab='Temp. Value')
plot(x[[7]],type='b',col='green',main="Change in Fluid Intake by Admission",xlab='Admission #', ylab='Fluid Intake Value')
}

plot(p)
```

3. Create a further class for a cohort (group) of patients, and write methods for `mean` and `print` which, when applied to a cohort, apply mean or print to each patient contained in the cohort. Hint: think of this as a "container" for patients.  Reset the RNG seed to 8 and create a cohort of ten patients, then show the output for `mean` and `print`. (5 points)
points)
```{r}
rm(.Random.seed, envir = globalenv())
set.seed(8)
cohort = list()
  for (i in 1:10){
    x = makePatient()
    names(x) <- c("name", "gender", "date_of_birth", "date_of_admission", "pulse", "temperature", "fluid_intake")
    class(x) <- 'medicalRecord'
    cohort[[i]] = x
}
class(cohort) <- 'cohort'
class(cohort)

#Mean Method for Cohort
mean.cohort <- function(x){
  lapply(x,mean)
}

mean(cohort)
```

```{r}
#Print Method for Cohort
print.cohort <- function(x){
  for (i in 1:10){
    print(cohort[[i]])
  }
}

print(cohort)
```